---
title: "Fig1_final"
output: html_document
date: "2026-08-18"
editor_options: 
  chunk_output_type: console
---

PURPOSE: Create Figure 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(collapse)
library(data.table)
library(sf)
library(tigris)
library(tmap)
library(tmaptools)
library(openxlsx)
library(ggh4x)
library(scales)
library(RColorBrewer)

options(scipen = 99999)

knitr::opts_chunk$set(echo = TRUE)
```


# Read in files
```{r}
folder = "ADD PATH HERE"

## Farmworker counts
file <- paste0(folder, "/01a_final data/01_farmworker_imputation_pred.rds")
df.farmworker <- fread(file) %>%
  mutate(variable = case_when(
    variable %in% c("Hired Labor", "Unpaid Labor") ~ "Hired+Unpaid Labor",
    variable == "Migrant Labor (incl hired)" ~ "Migrant Labor"
  )) %>%
  group_by(Year, STATEFIPS, STATEABB, STATENAME, CNTYFIPS, GEOID, region_2, region_1, variable) %>%
  summarise(across(c(n_farmworkers, n_farmworker_imput), ~sum(.))) %>%
  ungroup()


## Days of co-exposure to HI above thresholds
file = paste0(folder, "/01a_final data/03_HI_smoke_2008-2023.csv")
df.HI.smoke <- readRDS(file ) %>%
  mutate(region_2 = factor(region_2, levels = c(  "East North Central" , "Middle Atlantic",   "New England" ,
                                                 "West South Central", "East South Central" , "South Atlantic",
                                                 "Pacific", "Mountain", "West North Central" ))) %>%
  mutate(month.abb = fct_reorder(month.abb, month))

## Regions 
df.regions <- data_frame(STATEABB = state.abb) %>%
  mutate(region_2 = case_when(
              STATEABB %in% c("CT", "ME", "MA", "NH", "VT", "RI") ~ "New England",
              STATEABB %in% c("NY", "PA", "NJ") ~ "Middle Atlantic",
              STATEABB %in% c("OH", "IN", "IL", "MI", "WI") ~ "East North Central",
              STATEABB %in% c("DE", "MD", "DC", "VA", "WV", "NC", "SC", "GA", "FL") ~ "South Atlantic",
              STATEABB %in% c("KY", "TN", "AL", "MS") ~ "East South Central",
              STATEABB %in% c("TX", "LA", "AR", "OK") ~ "West South Central",
              STATEABB %in% c("ND", "MN", "SD", "IA", "NE", "KS", "MO") ~ "West North Central",
              STATEABB %in% c("MT", "WY", "ID", "NV", "UT", "CO", "AZ", "NM" ) ~ "Mountain",
              STATEABB %in% c("WA", "OR", "CA") ~ "Pacific"
              )) %>%
  filter(!STATEABB %in% c("AK", "HI"))

## Heat and WF smoke policies
file <- paste0(folder, "/Heat_WFsmoke_Policies.xlsx")
df.policies.og <- read.xlsx(file, sheet = "Sheet1")

df.policies <- df.policies.og %>%
  transmute(
    STATEABB = State,
    heat_OccStandard = case_when(
      Heat_ExistingOccRegulation == "Yes" ~ "Regulation Exists",
      Heat_Under.Development == "Yes" ~ "Regulation Under Development",
      str_detect(Heat_Active.Legislation, "Pending") ~ "Legistation Pending",
  ),
  WFSmoke_OccStandard = case_when(
    WFsmoke_ExistingOccRegulation == "Yes" ~ "Regulation Exists"
  )  ) %>%
  filter(!c(is.na(heat_OccStandard) & is.na(WFSmoke_OccStandard))) %>%
  full_join(., df.regions) %>%
  filter(!STATEABB == "MN") ## only indoor policies

## SF Counties
sf.counties <- counties(cb = TRUE, progress_bar = FALSE, year = 2021) %>%
  filter(!STUSPS %in% c("AK", 'HI', "PR", "AS", "MP", "GU", "VI")) %>%
#  filter(STUSPS %in% c("CT")) %>%
  transmute(STATEFIPS = as.numeric( STATEFP), CNTYNAME = NAME, STATEABB = STUSPS, GEOID = as.numeric(GEOID))  %>%
  full_join(., df.regions)

## SF States
sf.states <- sf.counties %>%
  group_by(STATEABB, region_2) %>%
  summarise(totalFIPS = fsum(STATEFIPS)) %>%
  ungroup() %>%
  st_make_valid()
  
## SF Regions
sf.regions <- sf.counties %>%
  group_by(region_2) %>%
  summarise(totalFIPS = fsum(STATEFIPS)) %>%
  ungroup() %>%
  st_make_valid()
```



# Figure 1A
Bar graph of change in average annual coexpos days above 80F threshold between 2019-2023 vs 2008-2012
```{r}
## Count annual co-exposure days above threshold
df.HI.nDays.annual <- df.HI.smoke %>%
  mutate(coexpos = if_else(smokePM_pred.wtd > 0 , "coexpos_days", "no coexpos"),
         coexpos = if_else(is.na(smokePM_pred.wtd), "no coexpos", coexpos)) %>%
  fmutate(indicator = case_when(
    Year %in% 2008:2012 ~ "2008-2012",
    Year %in% 2019:2023 ~ "2019-2023"
  )) %>% 
  select(region_2, region_1, GEOID, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME, Year, month, month.abb, yearmonday, crop_class, total_county_crop_area, mean.HI.wtd, smokePM_pred.wtd, coexpos, indicator) %>%
  fgroup_by(GEOID, region_2, region_1, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME,crop_class, coexpos, Year, indicator) %>%
  fsummarise(n_days_80F = fsum(mean.HI.wtd >=80   &  coexpos == "coexpos_days", na.rm = TRUE)
             ) %>%
  ungroup() %>%
  ## ADD POLICY LABELS
  full_join(., df.policies) %>%
  mutate(heat_WF_policies = case_when(
           !is.na(WFSmoke_OccStandard) ~ paste0("Heat Standard: ", heat_OccStandard, " & Wildfire Standard: ", WFSmoke_OccStandard),
           is.na(heat_OccStandard) ~ "Heat Standard: None",
           TRUE ~ paste0("Heat Standard: ", heat_OccStandard)
         ),
         heat_WF_policies = factor(heat_WF_policies, levels = c(
                  "Heat Standard: Regulation Exists & Wildfire Standard: Regulation Exists",
                      "Heat Standard: Regulation Exists",
                      "Heat Standard: Legistation Pending",
                      "Heat Standard: Regulation Under Development",
                      "Heat Standard: None"))
         )


## Calculate difference in average annual co-exposure days between 2019-2023 vs 2008-2012 periods
df.n.county.reg.1 <- df.HI.nDays.annual %>%
  filter(coexpos == "coexpos_days",
         crop_class == 1) %>%
  ### average annual co-exposure days
  group_by(GEOID, STATEABB, region_2, coexpos, crop_class, indicator, heat_WF_policies) %>%
  summarise(n.avg.days = mean(n_days_80F, na.rm = TRUE)) %>%
  select(indicator, STATEABB, region_2, GEOID, n.avg.days, heat_WF_policies) %>%
  ### Difference between 2019-2023 vs 2008-2012 in each county
  pivot_wider(names_from = indicator,
              values_from = n.avg.days) %>%
  mutate(thresholds = "n_days_80F",
         across(c(`2019-2023`, `2008-2012`), ~if_else(is.na(.), 0, .)),
         diff.avg.days = `2019-2023` - `2008-2012` ) %>%
  ungroup() %>%
  select(-`NA`) %>%
  ### Average county change in coexposure days by STATE
  group_by( region_2, STATEABB, coexpos, crop_class, thresholds, heat_WF_policies) %>%
  summarise(diff.avg.days = mean(diff.avg.days, na.rm = TRUE))  %>%
  ungroup()


## threshold labels
thresh = as_labeller(c(`n_days_80F`= "Average additional county\n co-exposure days \u226580\u00B0F") )

## assign policy colours
my_colors <- c("#fb8b24", "#e36414",  "#5f0f40", "#9a031e", "grey50")
names(my_colors) <- c("Heat Standard: Regulation Exists & Wildfire Standard: Regulation Exists",
                      "Heat Standard: Regulation Exists",
                      "Heat Standard: Legistation Pending",
                      "Heat Standard: Regulation Under Development",
                      "Heat Standard: None")

## Plot graph
x.1A <- df.n.county.reg.1 %>% 
  filter( thresholds == "n_days_80F") %>%
  mutate(label = case_when(
    diff.avg.days > 0 ~ paste0("+ ", round(diff.avg.days,1)),
    diff.avg.days < 0 ~ paste0("- ", round(diff.avg.days,1)),
  )) %>%
  ggplot() + 
  geom_bar(aes(x = diff.avg.days, y = reorder(STATEABB, diff.avg.days), fill = heat_WF_policies), stat = "identity", show.legend = FALSE ) +
  geom_text(aes(x = diff.avg.days, y = reorder(STATEABB, diff.avg.days), label = label), hjust = -0.05, size = 2.05) + 
  theme_bw() +
  facet_wrap(~thresholds, ncol = 1, labeller = thresh ) +
  scale_fill_manual(values = my_colors)+
  ylab("") +
  xlab("") +
  expand_limits(x = 85) +
  theme(aspect.ratio = 4,
        strip.background = element_rect(fill = "white", color = "black"),
        strip.text = element_text(size = 6),
        axis.title.x = element_text(size = 6),
        text = element_text(size = 6)
        )
```



# Figure 1B 
Map of EHSA hotspot clusters of days above 80F thresholds
  1. Read in ArcGIS EHSA results shp file
  2. Recategorize automatic EHSA patterns
        - New hotspot and Consecutive hotspots --> Recent hotspot
        - Persistent hotspot and Historical hotspot --> Consistent hotspot
        - Intensifying hotspot --> Intensifying hotspot
        - All others --> "No Pattern Detected"
  3. Label counties if it had a significant EHSA category was in 80F EHSA
  4. Outline states by Occupational Heat/Wildfire Standards
```{r}
## 80F
file.80 <- paste0(folder, "/01a_final data/EHSA_results/EHSA_natl_80F.shp") ### EHSA shp file
sf.80 <- st_read(file.80) %>%
    mutate(PATTERN = case_when(
    PATTERN %in% c("New Hot Spot", "Consecutive Hot Spot" ) ~ "Recent Hot Spot",
    PATTERN %in% c("Persistent Hot Spot", "Historical Hot Spot") ~ "Consistent Hot Spot",
    PATTERN %in% c("Sporadic Hot Spot", "Oscillating Hot Spot") ~ "No Pattern Detected",
    TRUE ~ PATTERN
  )) %>%
  transmute(LOC_LABEL, CATEGORY,
              PATTERN_80 = PATTERN)

sf.80 <- st_transform(sf.80, crs = st_crs(sf.counties))


## remove "no pattern detected" from 80F df
sf.80.c <- sf.80 %>%
  transmute(LOC_LABEL, CATEGORY, 
            THRESHOLD = "80F",
            PATTERN = PATTERN_80) %>%
  filter(!PATTERN %in% c("No Pattern Detected")) %>%
  filter(!is.na(PATTERN))  %>%
  left_join(., sf.counties %>% mutate(LOC_LABEL = paste0("G", GEOID)) %>% st_drop_geometry())

## generate hotspote palette
x <- c("#804d36",   "#9972af", "#af8e53")
palette.sel <- x
names(palette.sel) <- c("Consistent Hot Spot" , "Intensifying Hot Spot",   "Recent Hot Spot")
  

## add to county and state geometries to space out states/counties by region
##    - MD / DE are additionally separated to improve visualization
sf.counties.1 <- sf.counties %>%
  mutate(geometry = case_when(
    region_2 == "East North Central" ~ geometry + c(0.75,0.75),
    region_2 == "East South Central" ~ geometry + c(0.75,-0.5),
    region_2 == "Middle Atlantic" ~ geometry + c(1.5,1),
    region_2 == "Mountain" ~ geometry + c(-1.5,0.25),
    region_2 == "New England" ~ geometry + c(2,1.25),
    region_2 == "Pacific" ~ geometry + c(-3.5,0.5),
    region_2 == "South Atlantic" ~ geometry + c(3,-0.75),
    region_2 == "West North Central" ~ geometry + c(-0.5,0.5),
    region_2 == "West South Central" ~ geometry + c(-0.5,-0.75),
    TRUE ~ geometry + c(0,0)
  )) %>%
  mutate(geometry = case_when(
    STATEABB == "MD" ~ geometry + c(0.5, 0.6),
    STATEABB == "DE" ~ geometry + c(0.75, 0.75),
    TRUE ~ geometry + c(0,0)
  ))

sf.states.1 <- sf.states %>%
  mutate(  region_2 = case_when(
    STATEABB == "MD" ~ "MD",
    STATEABB == "DE" ~ "DE",
    TRUE ~ region_2
  ) ) %>%
  filter(!is.na(region_2)) %>%
  mutate(geometry = case_when(
    region_2 == "East North Central" ~ geometry + c(0.75,0.75),
    region_2 == "East South Central" ~ geometry + c(0.75,-0.5),
    region_2 == "Middle Atlantic" ~ geometry + c(1.5,1),
    region_2 == "Mountain" ~ geometry + c(-1.5,0.25),
    region_2 == "New England" ~ geometry + c(2,1.25),
    region_2 == "Pacific" ~ geometry + c(-3.5,0.5),
    region_2 == "South Atlantic" ~ geometry + c(3,-0.75),
    region_2 == "West North Central" ~ geometry + c(-0.5,0.5),
    region_2 == "West South Central" ~ geometry + c(-0.5,-0.75),
    TRUE ~ geometry + c(0,0)
  )) %>%
  mutate(geometry = case_when(
    STATEABB == "MD" ~ geometry + c(3.5, -0.15),
    STATEABB == "DE" ~ geometry + c(3.75, 0),
    TRUE ~ geometry + c(0,0)
  ) )


sf.regions.1 <- sf.states.1 %>%
  group_by(region_2) %>%
  summarise(totalFIPS = sum(totalFIPS)) %>%
  ungroup() 


sf.80.c.1 <- sf.80.c %>%
  mutate(geometry = case_when(
    region_2 == "East North Central" ~ geometry + c(0.75,0.75),
    region_2 == "East South Central" ~ geometry + c(0.75,-0.5),
    region_2 == "Middle Atlantic" ~ geometry + c(1.5,1),
    region_2 == "Mountain" ~ geometry + c(-1.5,0.25),
    region_2 == "New England" ~ geometry + c(2,1.25),
    region_2 == "Pacific" ~ geometry + c(-3.5,0.5),
    region_2 == "South Atlantic" ~ geometry + c(3,-0.75),
    region_2 == "West North Central" ~ geometry + c(-0.5,0.5),
    region_2 == "West South Central" ~ geometry + c(-0.5,-0.75),
    TRUE ~ geometry + c(0,0)
  )) %>%
  mutate(geometry = case_when(
    STATEABB == "MD" ~ geometry + c(0.5, 0.6),
    STATEABB == "DE" ~ geometry + c(0.75, 0.75),
    TRUE ~ geometry + c(0,0)
  ))



## policies - standardize the language and add geometry
sf.policies <- full_join(sf.states, df.policies) %>%
    mutate(heat_WF_policies = case_when(
             !is.na(WFSmoke_OccStandard) ~ paste0("Heat Standard: ", heat_OccStandard, " & Wildfire Standard: ", WFSmoke_OccStandard),
             is.na(heat_OccStandard) ~ "No Heat or Wildfire Standard",
             TRUE ~ paste0("Heat Standard: ", heat_OccStandard)
           ),
           heat_WF_policies = factor(heat_WF_policies, levels = c(
                    "Heat Standard: Regulation Exists & Wildfire Standard: Regulation Exists",
                        "Heat Standard: Regulation Exists",
                        "Heat Standard: Legistation Pending",
                        "Heat Standard: Regulation Under Development",
                        "Heat Standard: None"))
          ) %>%
    filter(!is.na(heat_OccStandard)) %>%
    select(-c(heat_OccStandard, WFSmoke_OccStandard)) %>%
    select(STATEABB, heat_WF_policies, geometry) 

sf.policies.1 <- sf.policies %>%
  full_join(., df.regions) %>%
  mutate(geometry = case_when(
    region_2 == "East North Central" ~ geometry + c(0.75,0.75),
    region_2 == "East South Central" ~ geometry + c(0.75,-0.5),
    region_2 == "Middle Atlantic" ~ geometry + c(1.5,1),
    region_2 == "Mountain" ~ geometry + c(-1.5,0.25),
    region_2 == "New England" ~ geometry + c(2,1.25),
    region_2 == "Pacific" ~ geometry + c(-3.5,0.5),
    region_2 == "South Atlantic" ~ geometry + c(3,-0.75),
    region_2 == "West North Central" ~ geometry + c(-0.5,0.5),
    region_2 == "West South Central" ~ geometry + c(-0.5,-0.75),
    TRUE ~ geometry + c(0,0)
  )) %>%
  relocate(geometry, .after = last_col()) %>%
  mutate(geometry = case_when(
    STATEABB == "MD" ~ geometry + c(0.5, 0.6),
    STATEABB == "DE" ~ geometry + c(0.75, 0.75),
    TRUE ~ geometry + c(0,0)
  ))

## generate palette for policies
my_colors <- c("#fb8b24", "#e36414",  "#5f0f40", "#9a031e", "grey50")
names(my_colors) <- c("Heat Standard: Regulation Exists & Wildfire Standard: Regulation Exists",
                        "Heat Standard: Regulation Exists",
                        "Heat Standard: Legistation Pending",
                        "Heat Standard: Regulation Under Development",
                        "Heat Standard: None")
  

## align / assign CRS
st_crs(sf.regions.1 ) <- st_crs(sf.regions)
st_crs(sf.counties.1 ) <- st_crs(sf.counties)
st_crs(sf.states.1 ) <- st_crs(sf.states)
st_crs(sf.80.c.1) <- st_crs(sf.80.c)
st_crs(sf.policies.1) <- st_crs(sf.policies)

# Plot EHSA map
x.80<- tm_shape(sf.counties.1) +
  tm_borders(col = "grey80",
             lwd = 0.1) +
  tm_shape(sf.80.c.1 ) +
  tm_polygons(fill = tm_vars( "PATTERN"),
              fill.scale = tm_scale_categorical(values = palette.sel),
              col = "grey80",
              lwd = 0.1) +
  tm_shape(sf.states.1) +
  tm_borders(col = "grey60",
             lwd = 0.5) +
  tm_shape(sf.regions.1) +
  tm_borders(col = "grey30",
             lwd = 0.75)  +
  tm_shape(sf.policies.1) +
  tm_borders(col = tm_vars("heat_WF_policies"),
             col.scale = tm_scale_categorical(values = my_colors),
             lwd = 1)


```



# Figure 1C
Bar graph of percent of counties with increase, decrease or no change in average migrant farmworker coexposures in teh 2019-2023 period, compared to 2008-2012 baseline.
```{r}
df.migrants <- df.farmworker %>%
  mutate(indicator = case_when(
    Year %in% 2008:2012 ~ "2008-2012",
    Year %in% 2019:2023 ~ "2019-2023"
  )) %>%
  # average number of migrant FW in each period
  fgroup_by(indicator, GEOID, STATEABB, STATENAME, region_2, variable) %>%
  fsummarise(mean_n_farmworker_imput = fmean(n_farmworker_imput)) %>%
  filter(!is.na(indicator)) %>%
  pivot_wider(names_from = indicator,
              values_from = mean_n_farmworker_imput) %>%
  # difference in migrant FW between periods
  mutate(diff.08.23 = `2019-2023` - `2008-2012`, 
         perc.08.23 = diff.08.23/`2008-2012`,
         perc.08.23.100 = diff.08.23*100/`2008-2012`) %>%
  filter(variable == "Migrant Labor") %>%
  select(GEOID, region_2, STATEABB, STATENAME, diff.08.23, perc.08.23) %>%
  filter(!STATEABB == "DC")


## Percent of counties with increase / decrease / no change in migrant FW co-exposure days
df.migrants.n <- df.migrants %>%
  group_by(region_2, STATEABB, STATENAME) %>%
  mutate(names = case_when(
    diff.08.23 > 0 ~ "Increase",
    diff.08.23 < 0 ~ "Decrease",
    diff.08.23 == 0 ~ "No Change or \nNo Migrant FW",
    is.na(diff.08.23) ~ "No Change or \nNo Migrant FW"
  )) %>%
  ungroup()  %>%
  mutate(names = factor(names, levels = c("Increase", "Decrease", "No Change or \nNo Migrant FW"))) %>%
  group_by(region_2, STATENAME, STATEABB) %>%
  mutate(n_counties = n()) %>%
  group_by(region_2, STATENAME, STATEABB, names, n_counties) %>%
  summarise(values = n()) %>%
  ungroup() %>%
  mutate(perc = values / n_counties)

## admin - organize states from highest to lowest percent of counties with increase
df.fct <- df.migrants.n  %>%
  filter(names =="Increase") %>%
  group_by(names) %>%
  arrange(names, -perc) %>%
  mutate(id_n = row_number()) %>%
  ungroup() %>%
  select(region_2, STATEABB, STATENAME, id_n) %>%
  ungroup()

## generate palette for increase / decrease / no change
my_colours = colorRampPalette(brewer.pal(3, "Dark2"))
my_colours <- my_colours(3)
names(my_colours) <- c("Increase", "Decrease","No Change or \nNo Migrant FW")

## plot graph
x <- df.migrants.n %>% 
  full_join(., df.fct) %>% 
  ## make sure that De, WV and RI are last
  mutate(
    id_n = case_when(
      STATEABB == "DE" ~ 46,
      STATEABB == "WV" ~ 47,
      STATEABB == "RI" ~ 48,
      TRUE ~ id_n
    )
  ) %>%
  mutate(STATEABB = fct_reorder(STATEABB, id_n, .desc = FALSE ),
         STATENAME = fct_reorder(STATENAME, id_n, .desc = FALSE)) %>%
  ggplot() +
  geom_bar(aes(x = STATENAME, y = values, fill = names), position = "fill", stat = "identity") +
  theme_bw() +
  theme(aspect.ratio = 0.25,
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 3, angle = 45, hjust = 0.95)
        ) +
  ylab("Percent of \ncounties") +
  scale_fill_manual(values = my_colours) +
  scale_y_continuous(labels = label_percent(),
                     breaks = seq(0,1, by = 0.25))

```



















