---
title: "10_Supplementary_Figures"
output: html_document
date: "2026-05-30"
editor_options: 
  chunk_output_type: console
---

PURPOSE: Create Supplementary Figures


```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(collapse)
library(sf)
library(tigris)
library(tmap)
library(tmaptools)
library(openxlsx)
library(ggh4x)
library(patchwork)
library(ggrepel)
library(scales)
library(svglite)
library(RColorBrewer)
library(ggview)

options(scipen = 99999)

knitr::opts_chunk$set(echo = TRUE)
```

# Read in files 
```{r}
folder = "ADD PATH HERE"

## Farmworkers 
file <- paste0(folder, "/01_data/01_farmworker_imputation_pred.csv")
df.farmworker <- fread(file) %>%
  mutate(variable = case_when(
    variable %in% c("Hired Labor", "Unpaid Labor") ~ "Hired+Unpaid Labor",
    variable == "Migrant Labor (incl hired)" ~ "Migrant Labor"
  )) %>%
  group_by(Year, STATEFIPS, STATEABB, STATENAME, CNTYFIPS, GEOID, region_2, region_1, variable) %>%
  summarise(across(c(n_farmworkers, n_farmworker_imput), ~sum(.))) %>%
  filter(Year %in% 2008:2023) 

## Days of exposure to HI above thresholds
file = "/home/jp3343/R_farmworker/03_outputs/v2_05a_HI_2008-2024.csv"
df.HI <- fread(file ) %>%
  mutate(region_2 = factor(region_2, levels = c(  "East North Central" , "Middle Atlantic",   "New England" ,
                                                 "West South Central", "East South Central" , "South Atlantic",
                                                 "Pacific", "Mountain", "West North Central" ))) %>%
  mutate(month.abb = fct_reorder(month.abb, month))


## Days of co-exposure to HI above threhsolds
file = "/home/jp3343/R_farmworker/03_outputs/v3_05c_HI_smoke_2008-2024.csv"
df.HI.smoke <- fread(file ) %>%
  mutate(region_2 = factor(region_2, levels = c(  "East North Central" , "Middle Atlantic",   "New England" ,
                                                 "West South Central", "East South Central" , "South Atlantic",
                                                 "Pacific", "Mountain", "West North Central" ))) %>%
  mutate(month.abb = fct_reorder(month.abb, month))


## Counties sf
sf.counties <- counties(cb = TRUE, progress_bar = FALSE, year = 2021) %>%
  filter(!STUSPS %in% c("AK", 'HI', "PR", "AS", "MP", "GU", "VI")) %>%
  transmute(STATEFIPS = as.numeric( STATEFP), CNTYNAME = NAME, STATEABB = STUSPS, GEOID = as.numeric(GEOID))

## state sf
sf.states <- sf.counties %>%
  group_by(STATEABB) %>%
  summarise(totalFIPS = fsum(STATEFIPS)) %>%
  ungroup() %>%
  st_make_valid()
  
## region sf
sf.regions <- sf.counties %>%
  mutate(region_2 = case_when(
    STATEABB %in% c("CT", "ME", "MA", "NH", "VT", "RI") ~ "New England",
    STATEABB %in% c("NY", "PA", "NJ") ~ "Middle Atlantic",
    STATEABB %in% c("OH", "IN", "IL", "MI", "WI") ~ "East North Central",
    STATEABB %in% c("DE", "MD", "DC", "VA", "WV", "NC", "SC", "GA", "FL") ~ "South Atlantic",
    STATEABB %in% c("KY", "TN", "AL", "MS") ~ "East South Central",
    STATEABB %in% c("TX", "LA", "AR", "OK") ~ "West South Central",
    STATEABB %in% c("ND", "MN", "SD", "IA", "NE", "KS", "MO") ~ "West North Central",
    STATEABB %in% c("MT", "WY", "ID", "NV", "UT", "CO", "AZ", "NM" ) ~ "Mountain",
    STATEABB %in% c("WA", "OR", "CA") ~ "Pacific"
  ) ) %>%
  group_by(region_2) %>%
  summarise(totalFIPS = fsum(STATEFIPS)) %>%
  ungroup() %>%
  st_make_valid()
```



# Supplementary Figure 1 - Change in average annual coexpos days above threshold between 2019-2023 vs 2008-2012
```{r}
## Count annual co-exposure days above threshold
df.HI.nDays.annual <- df.HI.smoke %>%
  mutate(coexpos = if_else(smokePM_pred.wtd > 0 , "coexpos_days", "no coexpos"),
         coexpos = if_else(is.na(smokePM_pred.wtd), "no coexpos", coexpos)) %>%
  fmutate(indicator = case_when(
    Year %in% 2008:2012 ~ "2008-2012",
    Year %in% 2019:2023 ~ "2019-2023"
  )) %>% 
  select(region_2, region_1, GEOID, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME, Year, month, month.abb, yearmonday, crop_class, total_county_crop_area, mean.HI.wtd, smokePM_pred.wtd, coexpos, indicator) %>%
  fgroup_by(GEOID, region_2, region_1, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME,crop_class, coexpos, Year, indicator) %>%
  fsummarise(n_days_80F = fsum(mean.HI.wtd >=80   &  coexpos == "coexpos_days"),
             n_days_90F = fsum(mean.HI.wtd >=90   &  coexpos == "coexpos_days"),
             n_days_100F = fsum(mean.HI.wtd >=100 & coexpos == "coexpos_days"),
             n_days_110F = fsum(mean.HI.wtd >=110 & coexpos == "coexpos_days")
             ) 

## Calculate difference in average annual co-exposure days between 2019-2023 vs 2008-2012 periods
df.n.county.reg.1 <- df.HI.nDays.annual %>%
  pivot_longer(names_to = "thresholds",
               values_to = "n Days >= threshold HI",
               cols = 12:15) %>%
  mutate(`thresholds` = factor(`thresholds`, 
                                           levels = c("n_days_80F",
                                                      "n_days_90F",
                                                      "n_days_100F",
                                                      "n_days_110F"))) %>%
  filter(coexpos == "coexpos_days",
         crop_class == 1) %>%
  ### average annual co-exposure days
  group_by(GEOID, region_2, coexpos, crop_class, indicator, thresholds) %>%
  summarise(n.avg.days = mean(`n Days >= threshold HI`)) %>%
  filter(thresholds %in% c("n_days_100F", "n_days_110F")) %>%
  select(indicator, region_2, GEOID, thresholds, n.avg.days) %>%
  ### Difference between 2019-2023 vs 2008-2012 in each county
  pivot_wider(names_from = indicator,
              values_from = n.avg.days) %>%
  mutate(diff.avg.days = `2019-2023` - `2008-2012` ) %>%
  ungroup() %>%
  select(-`NA`) %>%
  ### Average county change in coexposure days by region
  group_by( region_2, coexpos, crop_class, thresholds) %>%
  summarise(diff.avg.days = mean(diff.avg.days, na.rm = TRUE)) 


## assign thresholds colours
my_colours = colorRampPalette(brewer.pal(8, "Dark2"))(9)
thresh = as_labeller(c(`n_days_80F` ="Days \u226580\u00B0F",
              `n_days_90F`="Days \u226590\u00B0F",
              `n_days_100F`="Days \u2265100\u00B0F",
              `n_days_110F` ="Days \u2265100\u00B0F"))

## Plot graph
x <- df.n.county.reg.1 %>% 
  mutate(label = case_when(
    diff.avg.days > 0 ~ paste0("+ ", round(diff.avg.days,1)),
    diff.avg.days < 0 ~ paste0("- ", round(diff.avg.days,1)),
  )) %>%
  ggplot() + 
  geom_bar(aes(x = diff.avg.days, y = reorder(region_2, diff.avg.days), fill = region_2), stat = "identity", show.legend = FALSE ) + 
  geom_text(aes(x = diff.avg.days, y = reorder(region_2, diff.avg.days), label = label), hjust = -0.05, size = 2.05) + 
  theme_bw() +
  facet_wrap(~thresholds, ncol = 1, labeller = thresh) +
  scale_fill_manual(values = my_colours)+
  ylab("") +
  xlab("Average additional county co-exposure days") +
  expand_limits(x = 80) +
  theme(aspect.ratio = 0.7,
        strip.background = element_rect(fill = "white", color = "black"),
        strip.text = element_text(size = 6),
        axis.title.x = element_text(size = 6),
        text = element_text(size = 6)
        ) +
  canvas(width = 3.5, height = 5.5 ) #, units = "px")

```

Save
```{r}
filename = paste0(folder, "/03_outputs/SuppFig1_bar plot.png")
save_ggplot(x, filename)
```



# Supplementary Figure 2 - Line graph of total FW-HI exposure days and FW-co-exposure days, 2008-2023, 100F and 110F
## Calculate FW-days / FW-coexposure days
```{r}
## FW-HI exposure days
df.HI.FW <- df.HI %>%
  fgroup_by(GEOID, region_2, region_1, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME,crop_class,Year ) %>%
  fsummarise(n_days_80F = fsum(mean.HI.wtd >=80 , na.rm = TRUE),
             n_days_90F = fsum(mean.HI.wtd >=90 , na.rm = TRUE),
             n_days_100F = fsum(mean.HI.wtd >=100, na.rm = TRUE),
             n_days_110F = fsum(mean.HI.wtd >=110, na.rm = TRUE)
             ) %>%
  full_join(., df.farmworker) %>%
  pivot_longer(names_to = "thresholds",
               values_to = "n_days",
               cols = 10:13) %>%
  mutate(FW_days = n_days * n_farmworker_imput) %>%
  filter(crop_class == 1,) %>%
  select(GEOID, region_2, Year, crop_class, thresholds, FW_days, variable)


# FW-Coexpos days
df.HI.smoke.FW <- df.HI.smoke %>%
  mutate(coexpos = if_else(smokePM_pred.wtd > 0 , "coexpos_days", "no coexpos"),
         coexpos = if_else(is.na(smokePM_pred.wtd), "no coexpos", coexpos)) %>%
    fgroup_by(GEOID, region_2, region_1, STATEFIPS, STATEABB, CNTYFIPS, CNTYNAME,crop_class,Year, coexpos ) %>%
  fsummarise(n_days_80F = fsum(mean.HI.wtd >=80   &  coexpos == "coexpos_days"),
             n_days_90F = fsum(mean.HI.wtd >=90   &  coexpos == "coexpos_days"),
             n_days_100F = fsum(mean.HI.wtd >=100 & coexpos == "coexpos_days"),
             n_days_110F = fsum(mean.HI.wtd >=110 & coexpos == "coexpos_days")
             )  %>%
  full_join(., df.farmworker) %>%
  pivot_longer(names_to = "thresholds",
               values_to = "n_days",
               cols = 11:14) %>%
  mutate(FW_coexpos_days = n_days * n_farmworker_imput) %>%
  filter(coexpos == "coexpos_days",
         crop_class == 1,) %>%
  select(GEOID, region_2, Year, crop_class, thresholds, FW_coexpos_days, variable)

# Join df for graph and calculate national FW-days / FW-coexposure days
df.hi.FW.coexpos <- full_join(df.HI.FW, df.HI.smoke.FW) %>%
  filter(variable == "Hired+Unpaid Labor") %>%
  ## sum for national estimate
  group_by(Year, crop_class, thresholds, variable) %>%
  summarise(across(c(FW_days, FW_coexpos_days), ~sum(., na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(`thresholds` = factor(`thresholds`, 
                                           levels = c("n_days_80F",
                                                      "n_days_90F",
                                                      "n_days_100F",
                                                      "n_days_110F"))) %>%
  pivot_longer(names_to = "category",
               values_to = "FW_days",
               cols = 5:6) %>%
  mutate(category = case_when(
    category == "FW_days"  ~ "Farmworker-days",
    category == "FW_coexpos_days" ~ "Farmworker-days of coexposure"),
    category = factor(category, levels = c("Farmworker-days", "Farmworker-days of coexposure"))
  )

```


## Plot line graph
```{r}
x <- df.hi.FW.coexpos %>% 
  filter(thresholds %in% c("n_days_100F", "n_days_110F")) %>%
  ggplot() +
  geom_point(aes(x =Year, y= FW_days, colour = category), size = 0.5) +
  geom_line(aes(x =Year,  y= FW_days,  colour = category)) +
  theme_bw() +
    scale_y_continuous(
      name = "Farmworker-days (millions)",
      labels = label_number(scale = 1e-6 )  ) +
    scale_color_manual(values = c("#D95F02", "#1B9E77")) +
    facet_wrap(~thresholds, scales = "free", ncol = 1, labeller = thresh) +
  theme(aspect.ratio = 0.6) +
  theme(strip.background = element_rect(fill = "white", color = "black") ) +
    theme(legend.title = element_blank(),
          axis.title.x = element_blank(),
          text = element_text(size = 6, family = 'calibri')) +
  canvas(width = 6, height = 5.5 )

```

Save
```{r}
filename = paste0(folder, "/03_outputs/SuppFig2_line graph.png")
save_ggplot(x, filename)
```



# Supplementary Figure 3 - Bivariate map of EHSA hotspot clusters, 100F and 110F thresholds
Bivariate map of EHSA hotspot clusters of days above 100F / 110F thresholds
- Bivariate categories: EHSA hotspot category vs threshold (100F, 110F, both)
  1. Read in ArcGIS EHSA results shp file
  2. Recategorize automatic EHSA patterns
        - New hotspot and Consecutive hotspots --> Recent hotspot
        - Persistent hotspot and Historical hotspot --> Consistent hotspot
        - Intensifying hotspot --> Intensifying hotspot
        - All others --> "No Pattern Detected"
  3. Label counties if it had a significant EHSA category was in 100F EHSA, 110F EHSA, or in both
```{r}
## 100F
file.100 <- paste0(folder, "/01a_final data/EHSA_results/EHSA_natl_100F.shp") ### EHSA shp file
sf.100 <- st_read(file.100) %>%
    mutate(PATTERN = case_when(
    PATTERN %in% c("New Hot Spot", "Consecutive Hot Spot" ) ~ "Recent Hot Spot",
    PATTERN %in% c("Persistent Hot Spot", "Historical Hot Spot") ~ "Consistent Hot Spot",
    PATTERN %in% c("Sporadic Hot Spot", "Oscillating Hot Spot") ~ "No Pattern Detected",
    TRUE ~ PATTERN
  )) %>%
  transmute(LOC_LABEL, CATEGORY,
              PATTERN_100 = PATTERN)

## 110F
file.110 <- file.100 <- paste0(folder, "/01a_final data/EHSA_results/EHSA_natl_110F.shp") ### EHSA shp file
sf.110 <- st_read(file.110) %>%
    mutate(PATTERN = case_when(
    PATTERN %in% c("New Hot Spot", "Consecutive Hot Spot" ) ~ "Recent Hot Spot",
    PATTERN %in% c("Persistent Hot Spot", "Historical Hot Spot") ~ "Consistent Hot Spot",
    PATTERN %in% c("Sporadic Hot Spot", "Oscillating Hot Spot") ~ "No Pattern Detected",
    TRUE ~ PATTERN
  )) %>%
  transmute(LOC_LABEL, CATEGORY,PATTERN)

## Both 100F and 110F
df.110 <- st_read(file.110) %>%
    mutate(PATTERN = case_when(
    PATTERN %in% c("New Hot Spot", "Consecutive Hot Spot" ) ~ "Recent Hot Spot",
    PATTERN %in% c("Persistent Hot Spot", "Historical Hot Spot") ~ "Consistent Hot Spot",
    PATTERN %in% c("Sporadic Hot Spot", "Oscillating Hot Spot") ~ "No Pattern Detected",
    TRUE ~ PATTERN
  )) %>%
  st_drop_geometry() %>%
  transmute(LOC_LABEL, CATEGORY,
              PATTERN_110 = PATTERN)
  
### join EHSA 100F and 110F results; identify ones with the same pattern in both EHSA; filter for only those counties
sf.100.110 <- full_join(sf.100, df.110) %>%  
  mutate(PATTERN_BOTH = case_when(
    PATTERN_100 == PATTERN_110 ~ PATTERN_100
  )) %>%
  filter(!PATTERN_BOTH %in% c("No Pattern Detected")) %>%
  filter(!is.na(PATTERN_BOTH)) %>%
  mutate(THRESHOLD = "Both") %>%
  rename(PATTERN = PATTERN_BOTH) %>%
  select(LOC_LABEL, CATEGORY, THRESHOLD, PATTERN, geometry)


## remove "no pattern detected" from 100F and 110F df
sf.100.c <- sf.100 %>%
  transmute(LOC_LABEL, CATEGORY, 
            THRESHOLD = "100F",
            PATTERN = PATTERN_100) %>%
  filter(!PATTERN %in% c("No Pattern Detected")) %>%
  filter(!is.na(PATTERN))  %>%
  filter(!LOC_LABEL %in% unique(sf.100.110$LOC_LABEL))


sf.110.c <- sf.110 %>%
  transmute(LOC_LABEL, CATEGORY, 
            THRESHOLD = "110F",
            PATTERN ) %>%
  filter(!PATTERN %in% c("No Pattern Detected")) %>%
  filter(!is.na(PATTERN)) %>%
  filter(!LOC_LABEL %in% unique(sf.100.110$LOC_LABEL))
  

# Bind all df  
sf.all <- bind_rows(sf.100.110,
                    sf.100.c,
                    sf.110.c)
  
  
# Plot bivariate map
x.100.110<- tm_shape(sf.counties) +
  tm_borders(col = "grey80",
             lwd = 0.1) +
  tm_shape(sf.states) +
  tm_borders(col = "grey60",
             lwd = 0.5) +
  tm_shape(sf.regions) +
  tm_borders(col = "black",
             lwd = 0.1) +
  tm_shape(sf.all ) +
  tm_polygons(fill = tm_vars(c("THRESHOLD", "PATTERN"), multivariate = TRUE),
              fill.scale = tm_scale_bivariate(values = "purplegold"),
              fill.legend = tm_legend_bivariate(xlab = "",
                                                ylab = ""),
              lwd = 0.1)
```


Save map
```{r}
filename = paste0(folder, "/03_outputs/SuppFig3_map_100F110F.png")
tmap_save(x.100.110, filename)
```

